mathml: Translate R expressions to MathML and LaTeX/MathJax

### 2023-01-14 rj/mathml.Rmd
simplified hooks

hook(quote(mysign), quote(sg(x)))
term <- call("<-", quote(mysign), sgn)
mathout(term)
```

@@

```{r, results='asis'}
hook(quote(m_A), quote(mean(X)["A"])); hook(quote(s2_A), quote(s["A"]^2L));
hook(quote(n_A), quote(n["A"]))
hook(quote(m_B), quote(mean(X)["B"])); hook(quote(s2_B), quote(s["B"]^2L))
hook(quote(n_B), quote(n["B"])); hook(quote(s2_p), quote(s["pool"]^2L))

@@

hook(quote(successes), quote(k))
hook(quote(Ntotal), quote(N))
hook(quote(prob), quote(pi))
mathout(term)
```

To improve the situation, \pkg{mathml} provides a simple hook that can be used
to replace elements (e.g., verbose variable names) of the code by concise
mathematical symbols, as illustrated in the example. Care should be taken to
avoid recursive hooks such as `hook(quote(s), quote(s["A"]))` that endlessly
replace the \(s\) from \(s_{\mathrm{A}}\) as
in \(s_{\mathrm{A}_{\mathrm{A}_{\mathrm{A}\cdots}}}\).

@@

```{r, echo=FALSE}
hook(quote(mu_A), quote(mu["A"]))
hook(quote(mu_B), quote(mu["B"]))
hook(quote(sigma_A), quote(sigma["A"]))
hook(quote(sigma_B), quote(sigma["B"]))
hook(quote(mu_M), quote(mu["M"]))
hook(quote(zn), quote(z[tau]))
hook(quote(M), quote(overline(X)))
```

@@

```{r, results='asis'}
hook(call("ED_single", expression(c), expression(mu)), 
     call("E", call(";", quote(D), call(",", expression(c), expression(mu)))))

ED_single <- function(c, mu)
  dfrac(c, mu)

mathout(call("=", quote(ED_single(c, mu)), ED_single))
```

@@

```{r, results='asis'}
mathout(call("=", quote(E(D["AB"])), quote(ED_single(c, mu_A + mu_B))))
```

@@

```{r, results="asis"}
hook(call("ED_async", expression(tau), expression(c), expression(mu_A), expression(sigma_A), expression(mu_B)), 
     call("E", call(";", call("[", quote(D), expression(tau)), 
                    call(",", expression(c), expression(mu_A), expression(sigma_A), expression(mu_B)))))

ED_async <- function(tau, c, mu_A, sigma_A, mu_B)
{ dfrac(c, mu_A) + (dfrac(1L, mu_A) - dfrac(1L, mu_A + mu_B)) *
    ((mu_A*tau - c) * pnorm(dfrac(c - mu_A*tau, sqrt(sigma_A^2L*tau)))
      - (mu_A*tau + c) * exp(dfrac(2L*mu_A*c, sigma_A^2L))
        * pnorm(dfrac(-c - mu_A*tau, sqrt(sigma_A^2L*tau))))
}

mathout(call("=", quote(ED_async(tau, c, mu_A, sigma_A, mu_B)), ED_async))
```

@@

```{r, results='asis'}
hook(call("ED", expression(tau), expression(c), expression(mu_A), expression(sigma_A), expression(mu_B), expression(sigma_B)), 
     call("E", call(";", call("[", quote(D), expression(tau)), 
                    call(",", expression(c), expression(mu_A), expression(sigma_A), expression(mu_B), expression(sigma_B)))))

ED <- function(tau, c, mu_A, sigma_A, mu_B, sigma_B)
{
  if(tau == Inf) return(ED_single(c, mu_A)) ;
  if(tau == -Inf) return(ED_single(c, mu_B)) ;
  if(tau == 0L) return(ED_single(c, mu_A + mu_B)) ;
  if(tau > 0L) return(ED_async(tau, c, mu_A, sigma_A, mu_B)) ;
  if(tau < 0L) return(ED_async(abs(tau), c, mu_B, sigma_B, mu_A))
}

mathout(call("=", quote(ED(tau, c, mu_A, sigma_A, mu_B, sigma_B)), ED))
```

@@

hook(call("zv", expression(m), expression(s), expression(n), expression(tau), 
        expression(c), expression(mu_A), expression(sigma_A), expression(mu_B), expression(sigma_B), expression(mu_M)), quote(z[tau]))

@@

mathout(call("=", quote(X["8 df"]^2L), gof))
```

>>>>

hook(mysign, sg(x))
term <- call("<-", quote(mysign), sgn)
mathout(term)
```

@@

```{r, results='asis'}
hook(m_A, mean(X)["A"]) ; hook(s2_A, s["A"]^2L) ;
hook(n_A, n["A"])
hook(m_B, mean(X)["B"]) ; hook(s2_B, s["B"]^2L)
hook(n_B, n["B"]) ; hook(s2_p, s["pool"]^2L)

@@

hook(successes, k)
hook(Ntotal, N)
hook(prob, pi)
mathout(term)
```

To improve the situation, \pkg{mathml} provides a simple hook that can be used
to replace elements (e.g., verbose variable names) of the code by concise
mathematical symbols, as illustrated in the example. Care should be taken to
avoid recursive hooks such as `hook(s, s["A"])` that endlessly replace
the \(s\) from \(s_{\mathrm{A}}\) as
in \(s_{\mathrm{A}_{\mathrm{A}_{\mathrm{A}\cdots}}}\).

@@

@@

```{r, results='asis'}
ED_single <- function(c, mu)
  dfrac(c, mu)

# display as E(D; mu), c is a scaling parameter
hook(ED_single(.C, .Mu), E(`;`(D, .Mu)))
mathout(call("=", quote(ED_single(c, mu)), ED_single))
```

@@

```{r, results='asis'}
hook(mu_A, mu["A"])
hook(mu_B, mu["B"])
hook(sigma_A, sigma["A"])
hook(sigma_B, sigma["B"])
hook(mu_M, mu["M"])
hook(M, overline(X))

mathout(call("=", quote(E(D["AB"])), quote(ED_single(c, mu_A + mu_B))))
```

@@

```{r, results="asis"}
ED_async <- function(tau, c, mu_A, sigma_A, mu_B)
{ dfrac(c, mu_A) + (dfrac(1L, mu_A) - dfrac(1L, mu_A + mu_B)) *
    ((mu_A*tau - c) * pnorm(dfrac(c - mu_A*tau, sqrt(sigma_A^2L*tau)))
      - (mu_A*tau + c) * exp(dfrac(2L*mu_A*c, sigma_A^2L))
        * pnorm(dfrac(-c - mu_A*tau, sqrt(sigma_A^2L*tau))))
}

hook(ED_async(.Tau, .C, .MA, .SA, .MB), E(`;`(D[.Tau], `,`(.MA, .SA, .MB))))
mathout(call("=", quote(ED_async(tau, c, mu_A, sigma_A, mu_B)), ED_async))
```

@@

```{r, results='asis'}
ED <- function(tau, c, mu_A, sigma_A, mu_B, sigma_B)
{
  if(tau == Inf) return(ED_single(c, mu_A)) ;
  if(tau == -Inf) return(ED_single(c, mu_B)) ;
  if(tau == 0L) return(ED_single(c, mu_A + mu_B)) ;
  if(tau > 0L) return(ED_async(tau, c, mu_A, sigma_A, mu_B)) ;
  if(tau < 0L) return(ED_async(abs(tau), c, mu_B, sigma_B, mu_A))
}

hook(ED(.Tau, .C, .MA, .SA, .MB, .SB), E(`;`(D[.Tau], `,`(.MA, .SA, .MB, .SB))))
mathout(call("=", quote(ED(tau, c, mu_A, sigma_A, mu_B, sigma_B)), ED))
```

@@

@@

hook(zv(.M, .S, .N, .Tau, .C, .MA, .SA, .MB, .MM), z(.Tau))
mathout(call("=", quote(X["8 df"]^2L), gof))
```
